Poisson approximate likelihood compared to the particle filter

### 2024-09-15 ms.Rnw
recalculated table 1. changed to sbh and pfilter at mle with sbh init

@@

wwr_benchmark <- sum(apply(dat[,-1],2, function(x)
  arima(log(x + 1), order =c(2,0,1))$aic + 2 * sum(log(x+1))
))

>>>>

<<simulate-initial-for-palfilter>>=
n_sim <- 1000

sim_sbh_init_for_pal <- simulate(
  sir_sbh_init,
  nsim=n_sim,
  params=est_mle_vec_sbh,
  seed = 123456
)

s1_t1_vec <- c()
i1_t1_vec <- c()
r1_t1_vec <- c()
s2_t1_vec <- c()
i2_t1_vec <- c()
r2_t1_vec <- c()
s3_t1_vec <- c()
i3_t1_vec <- c()
r3_t1_vec <- c()

for(i in 1:n_sim){
  s1_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["S1",2]
  i1_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["I1",2]
  r1_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["R1",2]

  s2_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["S2",2]
  i2_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["I2",2]
  r2_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["R2",2]

  s3_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["S3",2]
  i3_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["I3",2]
  r3_t1_vec[i] <- sim_sbh_init_for_pal[[i]]@states["R3",2]
}

init_dist_sbh_mle <- c(
  S1=round(mean(s1_t1_vec),0),
  I1=round(mean(i1_t1_vec),0),
  R1=round(mean(r1_t1_vec),0),
  S2=round(mean(s2_t1_vec),0),
  I2=round(mean(i2_t1_vec),0),
  R2=round(mean(r2_t1_vec),0),
  S3=round(mean(s3_t1_vec),0),
  I3=round(mean(i3_t1_vec),0),
  R3=round(mean(r3_t1_vec),0)
)
@

<<palfilter-at-mle_sbh, echo=FALSE, message=FALSE, warning=FALSE>>=
prop <- 420

initial_guess_mif2_mle <- c(
  as.numeric(est_mle_vec_sbh["beta1"]), 
  as.numeric(est_mle_vec_sbh["beta2"]), 
  as.numeric(est_mle_vec_sbh["beta3"]),
  as.numeric(est_mle_vec_sbh["phi"]),
  as.numeric(est_mle_vec_sbh["beta11"]),
  as.numeric(est_mle_vec_sbh["sigma_q"]),
  as.numeric(est_mle_vec_sbh["sigma_xi"])
)

bake(file = paste0(out_dir,"pfilter/original_dataset/palfilter_mif2_mle.rds"),{
 set.seed(1223)
 rotavirus_SMC_qropxi(
    init_dist = init_dist_sbh_mle, 
    y_obs = realdat, 
    m = 9, 
    regular_params = c(
      initial_guess_mif2_mle[1], 
      initial_guess_mif2_mle[2], 
      initial_guess_mif2_mle[3], 
      initial_guess_mif2_mle[4],
      initial_guess_mif2_mle[5]
    ), 
    gamma_par = c(initial_guess_mif2_mle[7], 1/initial_guess_mif2_mle[7]), 
    norm_par = c(0.07, initial_guess_mif2_mle[6]), 
    prop = t(prop), 
    n_particles = NP_MIF, 
    ncores = ncores
  ) 
}) -> palfilter_mif2_mle
@

@@

wwr_benchmark <- sum(apply(realdat,2, function(x)
  arima(log(x + 1), order =c(2,0,1))$aic + 2 * sum(log(x+1))
))
